---
title: "chapter4.Rmd"
author: "Maria Karolina Wickholm"
date: "28 marraskuuta 2017"
output: html_document
---
<https://github.com/kkarolinaw/IODS-project-1>
```{r}
library(MASS)
```
```{r}
data("Boston")
```
```{r}
str(Boston)
dim(Boston)
```
The Boston data frame has 506 observations and 14 variables.
The data set consist of information needed to know while looking at housing values in suburbs in Boston.
The columns in the data set:
crim
per capita crime rate by town.
zn
proportion of residential land zoned for lots over 25,000 sq.ft.
indus
proportion of non-retail business acres per town.
chas
Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).
nox
nitrogen oxides concentration (parts per 10 million).
rm
average number of rooms per dwelling.
age
proportion of owner-occupied units built prior to 1940.
dis
weighted mean of distances to five Boston employment centres.
rad
index of accessibility to radial highways.
tax
full-value property-tax rate per \$10,000.
ptratio
pupil-teacher ratio by town.
black
1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.
lstat
lower status of the population (percent).
medv
median value of owner-occupied homes in \$1000s.
```{r}
pairs(Boston)
library(MASS)
library(tidyverse)
library(corrplot)
```
```{r}
cor_matrix<-cor(Boston) %>% round (digits = 2)
cor_matrix
corrplot(cor_matrix, method="circle", type = "upper", cl.pos = "b", tl.pos = "d", tl.cex = 0.6)
```
From the plot drawn with corrplot command, can be seen which variable correlates more or less with some other variables. Positive correlations are shown as blue and negative correlations as red. The bigger the dot the bigger the correlation. Here in the Boston data set can be seen that variables indus x dis (-0.71), nax x dis(-0.77) and age x dis(-0.75) have the largest negative correlations, and are shown as largest, most red pigmented dots. Variables rad x tax (0,91), indus x nox (0,76) and nox x age (0,73) all give pretty big, pigmented blue dots. Some variables have almost no correlation with each other, such are for example rm x chas (0,09). Looking at the chas variable, it's correlation is rather small with all othet variables too, variating from -0.12 to 0.18. Seems like the so called "Charles River dummy variable" has almost nothing or atleast just a little to do with Boston housing prizes.
```{r}
summary(Boston)
```

```{r}
boston_scaled <- scale(Boston)
summary(boston_scaled)
```
The scale command subtracts the column means from the corresponding columns and divide the difference with standard deviation. Here we can see the summary of Boston data set scaled. This is done so that linear discriminant analysis would be possible, since it requires the variables to have the same variance. 
```{r}
boston_scaled <- as.data.frame(boston_scaled)
```
```{r}
library(dplyr)
```
```{r}
summary(boston_scaled$crim)
bins <- quantile(boston_scaled$crim)
bins
crime <- cut(boston_scaled$crim, breaks = bins, include.lowest = TRUE, labels = c("low", "med_low", "med_high", "high"))
table(crime)
boston_scaled <- dplyr::select(boston_scaled, -crim)
boston_scaled <- data.frame(boston_scaled, crime)
n <- nrow(boston_scaled)
ind <- sample(n,  size = n * 0.8)
train <- boston_scaled[ind,]
test <- boston_scaled[-ind,]
correct_classes <- boston_scaled$crim
```

```{r}
lda.fit <- lda(crime ~ ., data = train)
lda.fit
lda.arrows <- function(x, myscale = 1, arrow_heads = 0.1, color = "red", tex = 0.75, choices = c(1,2)){
  heads <- coef(x)
  arrows(x0 = 0, y0 = 0, 
         x1 = myscale * heads[,choices[1]], 
         y1 = myscale * heads[,choices[2]], col=color, length = arrow_heads)
  text(myscale * heads[,choices], labels = row.names(heads), 
       cex = tex, col=color, pos=3)
}
classes <- as.numeric(train$crime)
plot(lda.fit, dimen = 2, col = classes, pch = classes)
lda.arrows(lda.fit, myscale = 1)
plot.new
```
Got the LD1 drawn WOOP finally!
#
problems with this code
test <- scale(test)
test_scaled <- as.data.frame(test)
lda.pred <- predict(lda.fit, newdata = test_scaled)
table(correct = correct_classes, predicted = lda.pred$class)

```{r}
data('Boston')
BostonAgain <- scale(Boston)
dist_eu <- dist(BostonAgain)
summary (dist_eu)
km <-kmeans(BostonAgain, centers = 3)
pairs(BostonAgain, col = km$cluster)
```
I think I'm having a little problem since it seems like the data measured with euklidean distance command would not let me print out the kmeans for the distanced data. I did however get BostonAgain data printed out by pairs command, but since this data set still lacks the distance measure, the plot it pretty impossible to read. I tried the as.data.frame function at first to create the data set including the measured distances, but somehow it didn't let me. If you have any suggestion how this could be fixed, would appreciate :)